A general and efficient representation of ancestral recombination graphs

### 2024-06-04 paper.tex
Update "probabilistic sampling" bit

Aside from these practical challenges, there is also a deeper
issue with the implicit strategy of basing an ARG data structure on
recording events and their properties (e.g.\ the crossover breakpoint
for a recombination event). This approach 
requires all events to be recorded explicitly, and does not 
provide an obvious mechanism for either aggregating multiple events
or expressing uncertainty about them. This is not a
problem when describing the results of simulations, where all details
are perfectly known. However, it can be an issue when we wish to
formally describe the output of various inference methods, particularly
those that avoid inferring events that are not \emph{knowable} from the data:
a useful approach as datasets approach the population scale~\citep[e.g.][]{
turnbull2018hundred, bycroft2018genome,hayes20191000,
Ros-Freixedes2020,karczewski2020mutational,tanjo2021practical,
halldorsson2022sequences}.

@@

% This is all a bit boring, maybe, but I think the people we want to
% influence into using tskit will need to have this spelled out for them.
Rigorously defining interchange formats~\cite[e.g.][]{kelleher2019htsget}
is difficult and time-consuming,
and no matter how precise the specification, in practise
it is the \emph{implementations} that determine how well methods
interoperate. The BAM read alignment format~\citep{li2009sequence}
is an instructive example. Originally developed as part of the
1000 Genomes project~\citep{10002015global} to address the fragmented
software ecosystem that existed at the time~\citep{danecek2021twelve},
BAM has since become ubiquitous in bioinformatics pipelines.
The excellent interoperability between methods
exchanging alignment data is largely attributable to the
success of \texttt{htslib}~\citep{bonfield2021htslib}, the
software library that \emph{implements} BAM and several other
foundational bioinformatics file formats.
Today, there are thousands of software projects
using \texttt{htslib}~\citep{bonfield2021htslib},
and it is this shared use of community software
infrastructure that guarantees the smooth flow of data between
applications.
The emerging ARG software ecosystem could similarly benefit
from the adoption of such shared community infrastructure
to handle the mundane and time-consuming details of data interchange.
The \texttt{tskit} library (Section~\ref{sec-efficiency})
is a high-quality open-source gARG implementation,
with proven efficiency and
scalability~\citep[e.g.][]{anderson2022genes,zhan2023towards},
that is already in widespread use.
Adopting it as a community standard may ease software implementation burden
on researchers, freeing their time to address the many fascinating open
questions and challenges that exist.

>>>>

Aside from these practical challenges, there is also a deeper
issue with the implicit strategy of basing an ARG data structure on
recording events and their properties (e.g.\ the crossover breakpoint
for a recombination event). This approach 
requires all events to be recorded explicitly, and does not 
provide an obvious mechanism for aggregating multiple, potentially
unresolvable, events.
As datasets approach the population scale~\citep[e.g.][]{
turnbull2018hundred, bycroft2018genome,hayes20191000,
Ros-Freixedes2020,karczewski2020mutational,tanjo2021practical,
halldorsson2022sequences} representing such uncertainty 
directly through the data structure is a useful alternative to 
classical methods based on probabilisitic sampling.

@@

% This is all a bit boring, maybe, but I think the people we want to
% influence into using tskit will need to have this spelled out for them.
Rigorously defining interchange formats~\cite[e.g.][]{kelleher2019htsget}
is difficult and time-consuming,
and no matter how precise the specification, in practise
it is the \emph{implementations} that determine how well methods
interoperate. The BAM read alignment format~\citep{li2009sequence}
is an instructive example. Originally developed as part of the
1000 Genomes project~\citep{10002015global} to address the fragmented
software ecosystem that existed at the time~\citep{danecek2021twelve},
BAM has since become ubiquitous in bioinformatics pipelines.
The excellent interoperability between methods
exchanging alignment data is largely attributable to the
success of \texttt{htslib}~\citep{bonfield2021htslib}, the
software library that \emph{implements} BAM and several other
foundational bioinformatics file formats.
Today, there are thousands of software projects
using \texttt{htslib}~\citep{bonfield2021htslib},
and it is this shared use of community software
infrastructure that guarantees the smooth flow of data between
applications.
The emerging ARG software ecosystem could similarly benefit
from the adoption of such shared community infrastructure
to handle the mundane and time-consuming details of data interchange.
The \texttt{tskit} library 
is a high-quality open-source gARG implementation,
with proven efficiency and
scalability~\citep[e.g.][]{anderson2022genes,zhan2023towards},
that is already in widespread use.
Adopting it as a community standard may ease software implementation burden
on researchers, freeing their time to address the many fascinating open
questions and challenges that exist.
